---
title: "TidyTuesday week 16: crossword puzzle"
author: "Christopher"
date: "4/20/2022"
output: html_document
---

# Reading and wrangling data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r}
library(tidyverse)
library(tidytext)
library(lubridate)
library(skimr)
theme_set(theme_minimal())

# Libraries
tuesdata <- tidytuesdayR::tt_load(2022, week = 16)
big_dave <- tuesdata$big_dave
times <- tuesdata$times

clues <- 
times %>% 
  bind_rows(big_dave) %>% 
  select(-rowid)
  
```

```{r}
clues %>% skim()
```



## What is the frequency of crossword puzzles by sources?

```{r}
clues %>%
  count(puzzle_date = floor_date(puzzle_date, "month"), source) %>% 
  ggplot(aes(puzzle_date, n, fill = source)) +
  geom_col()

```


## In which position does a definition located within the clue?

Definition can be found usually at the beginning and end of the clue.

```{r}
clues_def_pos <- 
clues %>% 
  extract(clue, c("clue", "lengths"), "(.*) \\((.*)\\)$") %>% 
  mutate(year = year(puzzle_date),
         definition_length = str_length(definition),
         definition_index = str_locate(clue, fixed(definition))[, 1],
         definition_position = (definition_index - 1) / (str_length(clue) - str_length(definition)))

clues_def_pos %>% 
  ggplot(aes(definition_position)) + 
  geom_histogram(boundary = 0, binwidth = 0.05)

```

## Percentage of missing crossword puzzle solutions across years by sources.

```{r}
clues_def_pos %>% 
  group_by(year, source) %>% 
  summarise(pct_missing = mean(is.na(definition_position))) %>% 
  ggplot(aes(year, pct_missing, color = source)) + 
  geom_line()
```


## What is the percentage that the definition will be found either start or end of the clue across available years?



```{r}
clues_def_pos %>% 
  filter(!is.na(definition_position)) %>%
  group_by(year, source) %>% 
  summarise(pct_start_or_end = mean(definition_position %in% c(0, 1)),
            n = n()) %>% 
  filter(n >= 1000) %>% 
  ggplot(aes(year, pct_start_or_end, color = source)) + 
  geom_line()
```

In the logistic regression, the percentage that definition will be at the start or end increases across year significantly, from 94% to 96%.

```{r}
clues_def_pos %>% 
  filter(!is.na(definition_position),
         source == "bigdave44") %>% 
  mutate(start_end = definition_position %in% c(0, 1)) %>% 
  glm(start_end ~ year, data = ., family = "binomial") %>% 
  summary()
```


## Do longer clues indicate longer answers?

Longer answers are associated with longer definition, but not by a lot (whether looking at the number of letters or number of words).

```{r}
clues_def_pos %>% 
  filter(!is.na(definition_position)) %>% 
  count(definition_length, answer_length = str_length(answer)) %>% 
  ggplot(aes(definition_length, answer_length, fill = n)) +
  geom_tile()
```

```{r}
clues_def_pos %>% 
  filter(!is.na(definition_position), !is.na(answer)) %>% 
  with(cor.test(definition_length, str_length(answer)))
```

```{r}
clues_def_pos_nwords <- 
clues_def_pos %>% 
  mutate(n_words_answer = str_count(lengths, ",") + 1,
         n_words_definition = str_count(definition, " ") + 1)

clues_def_pos_nwords %>% 
  filter(!is.na(definition), !is.na(lengths)) %>% 
  summarise(cor = cor(n_words_answer, n_words_definition))
```


Big Dave has been using shorter definitions over time as well as slightly more likely to be at the beginning or end.

```{r}
by_year_source <- 
clues_def_pos_nwords %>% 
  filter(!is.na(answer), !is.na(lengths)) %>% 
  group_by(year, source) %>% 
  summarise(n_puzzle = n(),
            avg_n_words_answer = mean(n_words_answer),
            avg_n_words_definition = mean(n_words_definition, na.rm = TRUE),
            n_clues_definitions = sum(!is.na(n_words_definition)),
            pct_start_end_or_end = mean(definition_position == 0 | definition_position == 1, na.rm = TRUE)) %>% 
  filter(n_clues_definitions >= 1000)
  
by_year_source %>% 
  ggplot(aes(year, avg_n_words_answer, color = source)) +
  geom_line() +
  labs(y = "Average number of words in the answer")

by_year_source %>% 
  ggplot(aes(year, avg_n_words_definition, color = source)) +
  geom_line() +
  geom_point(aes(size = n_clues_definitions)) +
  labs(y = "Average number of words in the definition")

by_year_source %>% 
  ggplot(aes(year, pct_start_end_or_end, color = source)) +
  geom_line() +
  labs(y = "% of time at start of end")

```



## Text analysis

```{r}

clue_words <- 
clues_def_pos_nwords %>% 
  filter(!is.na(clue)) %>% 
  mutate(clue = str_replace(clue, "’", "'")) %>% 
  unnest_tokens(word, clue)

clue_words %>% 
  count(word, sort = TRUE) %>% 
  anti_join(stop_words, by = "word") %>% 
  head(20) %>% 
  mutate(word = fct_reorder(word, n)) %>% 
  ggplot(aes(n, word)) + 
  geom_col()

```


Are the words used in cryptic crosswords different between the two sources?


























